\documentclass{article}

%
% Homework Details
%   - Title
%   - Due date
%   - Class
%   - Section/Time
%   - Instructor
%   - Author
%

\newcommand{\hmwkTitle}{AAA \ \#HW02}
\newcommand{\hmwkDueDate}{2022}
\newcommand{\hmwkClass}{Adv Abstract Algebra}
\newcommand{\hmwkClassTime}{Spr 2022}
\newcommand{\hmwkClassInstructor}{Prof. Peter Hermann}
\newcommand{\hmwkAuthorName}{\textbf{Xianzhi} }
\newcommand{\hmwkAuthor}{\textit{Xianzhi Wang}}

\input{preambleBloat.tex}
% \usepackage{xianzhi}

\begin{document}

\maketitle
Homework set 2 

\pagebreak

\begin{homeworkProblem}
    Let $\Gamma$ be an ordinary graph and denote the set of its 
    vertices by $V(\Gamma)$ and the set of its edges by $E(\Gamma)$.
    A bijection $\beta: V(\Gamma) \mapsto V(\Gamma)$ is
    called an \textit{automorphism} of $\Gamma$ if 
    for all $x,y \in V(\Gamma)$
    \begin{align}
        (x,y) \in E(\Gamma) \iff (\beta(x), \beta(y)) \in E(\Gamma)
    \end{align}
    holds. Then $Aut \Gamma := \left\{ \beta | \beta \ \text{is an automorphism of} \ \Gamma \right\}$
    is a group w.r.t. composition. 
    (It is called the \textit{automorphism group of} $\Gamma$. 
    It obviously acts on $V(\Gamma)$.)
    \begin{enumerate}
        \item Consider the following graph (the Petersen graph) below.
            Show that its automorphism group acts transitively
            on the $10$ vertices, i.e., all vertices belong to the same orbit.\\
        \item Determine the order of the automorphism group of the Petersen graph.
    \end{enumerate}
    

    


    
\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
    A transparent round disc is divided into five congruent sectors.
    Each sector is painted red or green or white.
    In how many ways can it be done if two such colorations count
    the same if and only if one can get them by means of a symmetry of the disc
    (a rotation or a reflection) from each other?

    \solution
    
    We have 
    \begin{align}
        \# \text{orbits} &= \frac{ 1 }{ \lvert G \rvert  }\sum_{g \in G} \# \ \text{of fixed points of} \ g
    \end{align}
    Since $2$ colorations are counted the same
    if they belong to the same orbit under $D_5$,
    we want to know the number of orbits.\\
    Rotation:
    \begin{align}
        &Id: \ \text{fix} \ 243 \ \text{elements}\\
        &a^1: 3\\
        &a^2: 3\\
        &a^3: 3\\
        &a^4: 3\\
    \end{align}
    Reflection:\\
    All reflections fix $3 \times 3 \times 3 = 27$ elements
    \begin{align}
        \# \text{orbits} &= \frac{ 1 }{ \lvert G \rvert  }\sum_{g \in D_5} \# \ \text{of fixed points of} \ g \\
        &= \frac{ 1 }{ 10 } (243 + 4 \cdot 3 + 5 \cdot 27)\\
        &= \frac{ 1 }{ 10 } (135 + 12 + 243) \\
        &= 39.
    \end{align}


    

    
    
\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
     Let $G$ be a finite group and 
     $k$ an integer. 
     Prove that $\left\{ g^k | g \in G \right\} = G$
     if and only if $gcd \left( \lvert G \rvert, k \right) = 1$.\\
 
     \solution
 
     $\left\{ g^k | g \in G \right\} = G \iff gcd( \lvert G \rvert, k ) = 1$.\\
     We show ``$\implies$''\\
     Assume for contradiction $gcd ( \lvert G \rvert, k) \neq 1$,
     so there exist a smallest prime divisor $p$ 
     such that $p \mid \lvert G \rvert,$ and $p \mid k$.\\
     By sylow's theorems, there exist $h \in G$ such that
     $\lvert h \rvert = p$. Thus,\\
     \begin{align}
         \lvert h^k \rvert &= \frac{ p }{ gcd(k,p) } = \frac{ p }{ p } = 1\\
         \implies h^k &= e.
     \end{align}
     Now, since $e^k = e$,
     \begin{align}
         \left\{ g^k \mid g \in G \right\}
     \end{align}
     has cardinality strictly less than $G$,\\
     so $\left\{ g^k | g \in G \right\} \neq G$, contradiction.\\

     We show "$\Leftarrow$" in 
     \begin{align}
         \left\{ g^k | g \in G \right\} = G &\iff gcd( \lvert G \rvert, k) = 1\\
     \end{align}
     Since $g \in G$, we have $g^k \in G \ \forall \ g$.
     \begin{align}
        \left\{ g^k \mid g \in G  \right\} \subset G
     \end{align}
     follows.\\
     Let $y \in G$. Use Euclidean Algorithm 
     $\exists x_1, x_2 \in \mathbb{Z}$.
     \begin{align}
         y &= y' = y^{gcd (\lvert G \rvert, k)} \\
         &= y^{x_1 \lvert G \rvert + x_2 k}\\
         &=y^{x_1 \lvert G \rvert} y^{x_2 k}\\
         &= y^{x_2 k}\\
         &=(y^{x_2})^k
     \end{align}
     Thus, $\forall \ y \in G$,
     we could express $y$ as $y^{x_2}$
     raise to the power $k$. So
     \begin{align}
         \left\{ g^k \mid g \in G \right\} &\supseteq G\\
         \implies \left\{ g^k \mid g \in G \right\} &= G
     \end{align}
     

     
     
     
     
    


    
\end{homeworkProblem}

\pagebreak

\end{document}







